---
title: "Control Structures"
output: html_document
---

Control structures are divided into 2 types :

1. Excute certain commands only when certain condition(s) is satisfied.</br>
  ***For e.g :*** *`if-then-else`* statements.
2. Execute certain commands repeatedly and use a certain logic to stop the iteration.</br> ***For e.g :*** *`for loop`*, *`while loop`* etc.


### `If else` Family

`if`, `if-else` and `if-elseif-else` are a family of constructs where "

- A condition is first checked, if satisfied then the operations are performed.
- If the condition is not satisfied, then, code exits the construct or, moves on to other options.

#### 1. `If` Construct

***Syntax :***
```{r if Syntax, eval= FALSE}
if(condition){
  statements..
}
```
#### 2. `If-else` Construct

***Syntax :***
```{r if-else Syntax, eval= FALSE}
if(condition){
  statements
} else{
  alternate statements
}
```

#### 3. `If-elseif-else` Construct

***Syntax :***
```{r if-elseif-else Syntax, eval= FALSE}
if(condition){
  statements
} else if(condition){
  alternate statements
} else{
  alternate statements
}
```

***Example :***

<i>Let's write the R-code for the following problem :</br>
- If `x` is greater than `7`,then, operations within the curved braces will occur.</br>
- Else the next condition,i.e., `x>8` would be checked and if this is not true</br>
- Final else condition is checked and if that too is false, then, there is no change is `x`.</i>

Let's assume the initial value of x=6 . So, the code would be :

```{r if-else example}

x = 6
if(x>7){
  x = x+1
}else if(x>8){
  x=x+2
}else{
  x = x+3
}

print(x)
```

### For Loop Construct

The structure of a `for loop` construct comprises :

- A sequence which could be a vector or, list.
- `iter` is an element of the sequence
- statements

Before understanding the *for loop*, we first need to understand the *Sequence function*.

#### Sequence Function

- A sequence function is one of the components of *for loop*.
- ***Syntax :***
```{r Sequence Function Syntax, eval = FALSE}
seq(from, to, length)
```
- The sequence function creates an equi-spaced points between `from` and `to`
  - `from` : Starting Number
  - `to` : Ending Number
  - `by` : Increment/Decrement (width)
  - `length` : Number of elements required

- ***Example :***

```{r sequence example}

seq(from = 1, to = 10, by = 2)

seq(from = 1, to = 10, length = 4)
```

#### For Loop
***Syntax :***
```{r for loop Syntax, eval = FALSE}

for(iter in sequence){
  statements
}

```

#### Nested For Loop

When one or, more `for loop` constructs are located within one another then, its known as *"Nested For Loop"*

***Syntax :***
```{r nested for loop Syntax, eval = FALSE}

for(iter_1 in sequence_1){
  for(iter_2 in sequence_2){
    statements
  }
}

```

***Example :***

```{r for loop example}

n = 5
sum = 0

for(i in seq(1,n,1)){
  sum = sum + i
  print(c(i,sum))
}

```

#### For Loop with `if-break`

- A `break` statement once executed, program exits the loop even before the iterations are complete.
- `break` command comes out of the innermost loop for nested loops.

***Example :***
*Let's say we want to terminate the looping when `sum` is greater than `15` :*

```{r for loop with break example}

n = 100
sum = 0

for (i in seq(1,n,1)){
  sum = sum + i
  print(c(i, sum))
  
  if(sum >= 15){
    break
  }
}
```

### While Loop

A `while loop` is used whenever we want to execute statements until a specific condition is violated.

***Example :***
<i> Consider a sequence of natural numbers. </br>

What is the value of the natural number up to which the calculated sum is less than the specified `Fin_sum` ?</i>

***1+2+3+.....+n = Fin_sum(15)***

```{r while loop example}

sum = 0
i = 0
Fin_sum = 15

while(sum < Fin_sum){
  i = i+1
  sum = sum+i
  print(c(i,sum))
}
```










