---
title: "Data Visualization In Basic R Graphics"
output: html_document
---
With the base R package, we can produce various plots to visualize our data but, the charts are less pleasing and we have to write complex R-codes too.

So, for data visualization purpose, we mostly use the *"ggplot2"* package, that has sophisticated graphics.

Still, let's see how we can generate some simple charts using the basic R graphics.

### Scatter Plot

Here is an example of a basic scatter plot in R :

```{r scatter plot}

X = 1:10
Y = X^2

plot(Y)
```

#### Analysing The *mtcars* Dataset

- *mtcars* is a dataframe with 32 observations on 11 variables and comes with the base-R package.

- The data was extracted from the *1974-Motor Trend US Magazine*, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles(1973-74 models).

Let's have a summary view of *mtcars* dataset before plotting any chart :

```{r mtcars summary}
summary(mtcars)
```

Now, let's see a corelation between the `wt` (Car Weight) & `mpg` (Miles Per Gallon) :

```{r mtcars scatter}
attach(mtcars)

plot(wt,mpg,
     main = "Car Weight ~ MPG Scatter",
     xlab = "Car Weight",
     ylab = "Miles Per Gallon",
     pch = 19)
```

### Line Plot

Here is an example of a basic line plot in R :

```{r Line plot}

X = 1:10
Y = cos(X)

plot(X,Y,
     type = "l")
```

### Bar Plot

Here is an example of a basic line plot in R :

```{r barplot}

H = c(7, 12, 28, 5, 41)

M = c("Mar", "Apr", "May", "Jun", "Jul")

barplot(H, names.arg = M,
        xlab = "Month",
        ylab = "Revenue",
        main = "Revenue Chart",
        col = "darkseagreen",
        border = "aliceblue"
        )
```




