---
title: "Dataframes In R"
author: "Udit Kumar Chatterjee"
output: html_document
---

### Create A Dataframe

Let's create a dataframe using 3 vectors :
```{r}
vec1 = c(1,2,3)
vec2 = c("R","Python","Java")
vec3 = c("For Prototyping", "For Prototyping", "For Scaleup")

df = data.frame(vec1,vec2,vec3)
print(df)
```

### Accessing Elements From A Dataframe

- To access the 1st and 2nd row of the dataframe `df`, we have to write the following code :

```{r}
print(df[1:2,])
```

- To access the 1st and 2nd column of the dataframe `df`, we have to write the following code :

```{r, eval= FALSE}
print(df[,1:2])
```
or,
```{r}
print(df[1:2])
```
### Extracting Subset From A Dataframe

Let's create another dataframe :

```{r}
pd = data.frame("Name"= c("Senthil","Senthil","Sam","Sam"),
                "Month"= c("Jan","Feb","Jan","Feb"),
                "BS"= c(141.2,139.3,135.2,160.1),
                "BP"= c(90,78,80,81))
print(pd)
```

The function `subset()`extracts the subset of the dataframe based on certain conditions given.

```{r}
pd2 = subset(pd,Name = "Senthil", BS>150)
print(pd2)
```

### Editing A Dataframe


Let's Change the value of 2nd column and 2nd row of dataframe `df` to "R" :

```{r}
df[[2]][2]="R"
print(df)
```
or, we can directly do this from a GUI by using the following command
```{r, eval= FALSE}
edit(df)
```

### GUI To Create & Edit Dataframe

We can create and edit the dataframes from GUI by using the following simple codes :

- For creating an empty dataframe :

```{r,eval=FALSE}
MyTable = data.frame()
```
- For adding data to the above created dataframe :

```{r, eval= FALSE}
MyTable = edit(MyTable)
```
- To view the created dataframe :
```{r, eval=FALSE}
View(MyTable)
```

### Adding Extra Rows & Columns To Dataframe

- For adding an extra row, we can use `rbind()` function :

```{r}
df = rbind(df,data.frame(vec1 = 4,
                         vec2 = "C",
                         vec3 = "For Scaleup"))
print(df)
```
- For adding an extra column, we can use `cbind()` function :

```{r}
df = cbind(df, vec4 = c(10,20,30,40))
print(df)
```

We can also do the same operations from a GUI by using the `edit()` function as discussed earlier.

### Deleting Rows & Columns From Dataframe

Let's create another dataframe in which we will show the data of `df` except the data of 3rd row and 1st column :

```{r}
df2 = df[-3,-1]
print(df2)
```

So, just by puttinge a negative sign (-), we can remove the desired entries of a dataframe.

#### Conditional Deletion

- To delete the 3rd column :

```{r}
df3 = df[,!names(df) %in% c("vec3")]
print(df3)
```

- To delete the 3rd row :

```{r}
df4 = df[!df$vec1==3,]
print(df4)
```

### Manipulating Rows - The Factor Issue

- When character columns are created in a data frame, they became factors.
- `df[3,1] = 3.1` command changes the value of 3rd row and 1st column to `3.1` in the dataframe `df` and also changes the datatype of `Vec1` from Integer datatype to decimal. But, `df[3,3] = "Others"` command neither changes the value of  3rd column and 3rd row to `Others` nor it changes the datatype of `Vec3` (A character column) but, rather returns an error.
- To avoid such factor issue of character columns, we have to pass an additional argument in the data frame command, as follows -:

```{r}
df=data.frame(vec1,vec2,vec3,stringsAsFactors = FALSE)
```
- By default, the `stringAsFactors` argument is `TRUE` and that's why it takes a character column as a factor.
- After passing the `stringAsFactor` argument as `FALSE`, if we command  `df[3,3] = "Others"` then, the value of 3rd column and 3rd row will be changed to `Others`.











