---
title: "Functions"
author: "Udit Kumar Chatterjee"
output: html_document
---

- A function accepts input arguments and produces output by executing valid R commands present in the function.
- Function name and file names need not to be the same.
- A file can have one or, more function definitions.
- Functions are created using the command `function()`

Syntax :

```{r, eval = FALSE}
f = function(arguments){
      statements
      return(argument)
    }
```

***Example :***

- We have already created a R-script file that consists of various custom function like `volcylinder()` , `volsacylinder_mimo()` etc. Now, in order to apply them, we need to call and load the function file and then use it.

- Let's calculate the volume of a cylinder that has a diameter of 15 units and lenggth of 30 units, by using the custom function `volcylinder()`


```{r}
source("Custom Functions.R") #Invoking and loading the custom functions
v = volcylinder(15,30)
v
```

### Passing Arguments To Functions

We can pass arguments to the function in the following ways :

- Passing the arguments in the same order as in the function definition.
```{r}
v = volcylinder(15,30) #Here, dia = 15 & len = 30
v
```
- Passing the arguments in different order by specifying the argument name.
```{r}
v = volcylinder(len = 30, dia = 15)
v
```
- Default values are used by the function if some or, all arguments are not passed.
```{r}
v = volcylinder() #Here, dia = 5 & len = 10 has been taken as arguments
v
```

### Lazy Function Evaluation

- Functions are lazily evaluated, which means that if some arguments are missing, then, the function is still executed as long as the execution doesn't involve these arguments

***Example :***

```{r}
volume_case1 = function(dia, len, rad){
  vol = pi * dia/2 * len/4
  return(vol)
}

vol = volume_case1(len=22, dia=12)
vol
```
In the above case, even if we didn't gave any value for the `rad` argument, the function executed correctly as the `rad` argument is not used anywhere.

But, in the below example; we will get an error because, the `rad` argument has been used in the function but, missing in our evaluation statement.

```{r, eval= FALSE}
volume_case2 = function(dia, len, rad){
  vol = pi * dia/2 * len/4
  print(rad)
  return(vol)
}

vol = volume_case2(len=22, dia=12)
vol
```

### MIMO Functions

- Functions in R can take multiple input objects but, returns only one object as output.
- However, this is not a limitation because, list objects can also be returns by a function.

***Example :***

- We have created a custom function called `volsacylinder_mimo()` to calculate both volume and surface area of a cylinder by taking diameter and length of the cylinder as input.

```{r}
source("Custom Functions.R")
result = volsacylinder_mimo(dia = 12, len = 15)

result["Volume"] #Getting only volume
result["Surface Area"] #Getting only surface area
```

### Inline Functions

- Inline functions comes very handy when we have to find out the output of a complex polynomial equation for various input values

***For Example :***

```{r}
func = function(x){x^2+4*log(x)+5*x+2}
```
```{r}
func(5)
```
```{r}
func(13)
```
```{r}
func(23)
```

### Looping Over Objects

There are few looping functions available in R that are pretty useful when working interactively on a command line.</br>

Those functions are as follows :

- `apply()` : Apply a function over the margins of an array or, matrix.
- `lapply()` : Apply a function over a list or, vector.
- `tapply()` : Apply a function over a ragged array.
- `mapply()` : Multivariate version of `lapply()`.
- `xxply()` : Function from `plyr` package.


#### 1. `apply()` Function

- Applies a given function over the margins of given array.
- ***Syntax :***
  ```{r, eval = F}
  apply(array, margins, function, ...)
  ```
- Here, `margins` refer to the dimension of the array along which the function need to be applied.
  - For rows, `margin = 1`
  - For columns, `margin = 2`
  
***Example :***
```{r}

K = matrix(1:9,3,3)
K
```
```{r}

apply(K,1,sum) #applying the sum function across the rows
```
```{r}
apply(K,2,sum) #applying the sum function across the columns
```

#### 2. `lapply()` Function

- Used to apply a function over a list.
- Always returns a list of the same length as the input list.
- ***Syntax :***
  ```{r, eval = F}
  lapply(list, function, ...)
  ```
***Example :***

*Let's take 2 matrices as the elements of a list and find their determinant in the output using in-built `det` function*

```{r}
R = matrix(1:9,3,3)
S = matrix(10:18,3,3)

MyList = list(R,S)

determinant_list = lapply(MyList, det)

determinant_list
```

#### 3. `mapply()` Function

- Multivariate version of `lapply()`.
- A function can be applied over several lists simultaneously.
- ***Syntax :***
  ```{r, eval = F}
  mapply(function, list1,list2,list3 ...)
  ```
***Example :***

*Let's apply the custom function `volcylinder()` over a set of diameters and lengths*

```{r}
source("Custom Functions.R")

dia = list(1,2,3,4,5,6)
len = list(7,8,9,10,11,12)

vol = mapply(volcylinder,dia,len)

vol
```

#### 4. `tapply()` Function

- Used to apply a function over subset of vectors given by a combination of factors.
- ***Syntax :***
  ```{r, eval = F}
  tapply(vector, factors, function ...)
  ```
***Example :***

*Let's apply the sum function to add the `VALUES` by `ID`*

```{r}

ID = c(1,1,1,1,2,2,2,3,3)
VALUES = c(1:9)

tapply(VALUES, ID, sum)
```












