---
title: "Matrix Operations In R"
View : 
output: html_document
---

## Matrix

Matrix is a rectangular arrangement of numbers in rows and columns.

In a matrix, rows run horizontally and columns run vertically.

### Creating Matrices

#### 1. General Way To Create A Matrix

```{r}
A = matrix( c(1,2,3,4,5,6,7,8,9),
            nrow = 3,
            ncol = 3,
            byrow = TRUE )
print(A)
```

- The first argument provides the input for matrix.
- The second argument provides the number of rows.
- The third argument provides the number of columns.
- `byrow = TRUE` means the elemnts of matrix will be filled row-wise.
- By default, `byrow = FALSE`, so the elements are arranged column-wise first.

#### 2. Matrix with Single Constant Value

If we have to create a matrix where all the rows and columns are to be filled with a single constant element then we can use the following code :

```{r}
B = matrix(3,4,5)
print(B)
```

#### 3. Diagonal Matrix

In a diagonal matrix; except the primary diagonal, all other values are null.

We can create a diagonal matrix as follows :

```{r}
k = c(4,5,6)
C = diag(k,3,3)
print(C)
```

#### 4. Identity Matrix

An identity matrix is a diaagonal matrix with all the diagonal element as 1.

```{r}
I = diag(1,3,3)
print(I)
```

### Exploring A Matrix

To find the size of a matrix :

```{r}
dim(A)
```
To find the number of rows in a matrix :

```{r}
nrow(A)
```
To find the number of columns in a matrix :

```{r}
ncol(A)
```
To find the length of the matrix, i.e., total number of elements in a matrix :

```{r, eval = FALSE}
length(A)
```

or,

```{r}
prod(dim(A))
```

### Accessing Matrix Elements

Strings can be assigned as the names of rows and columns using `rownames()` and `colnames()` functions.

Let's create a $(3\times3)$ matrix $M$ with appropriate row and columns names :

```{r}
elements = c(10,20,30,40,50,60,70,80,90)

M = matrix(elements, 3, 3, byrow = T)
colnames(M) = c("a", "b", "c")
rownames(M) = c("d","e", "f")

print(M)
```
To access the 1st and 2nd column of the matrix :

```{r}
M[,1:2]
```

To access the 1st and 3rd columns using the assigned column names :

```{r}
M[,c("a", "c")]
```

To access the 1st and 3rd rows using the assigned row names :

```{r}
M[c("d","f"),]
```

To access the element of 2nd row and 3rd column :

```{r}
M[2,3]
```
To get all the elements of the 1st column :

```{r}
M[,1]
```
To get all the elements of the 2nd row :

```{r}
M[2,]
```
To access all the elements of the matrix except the 2nd column :

```{r}
M[,-2]
```

To access all the elements of the matrix except the 2nd row :

```{r}
M[-2,]
```

### Colon Operator & Sub-Matrix

Colon operator can be used to create a row matrix :

```{r}
1:10 #List of numbers from 1 to 10
10:1 #List of numbers from 10 to 1
```

Create a sub-matrix from matrix $M$ that has first 3 rows and first two columns :

```{r, eval= FALSE}
M[1:3,1:2]
```
Or,

```{r, eval= FALSE}
M[1:3,-3]
```
Or,
```{r}
M[,1:2]
```

Create a sub-matrix from matrix $M$ that has 1st and 3rd row elements of 1st 2 columns :

```{r, eval= FALSE}
M[c(1,3),1:2]
```
or,
```{r}
M[c(1,3),c(1,2)]
```

### Matrix Concatenation

Matrix concatenation refers to merging of row or, column to a matrix.

- Concatenation of a row to a matix is done using `rbind()`.
- Concatenation of a column to a matrix is done using `cbind()`.

Consistency of the dimensions between the matrix and the vectors should be checked before concatenation.

Let's add an additional row to the matrix $A$ :

```{r}
print(A)
```
```{r}
X = matrix(c(10,11,12),1,3)
print(X)
```
```{r}
Y = rbind(A,X)
print(Y)
```

Similarly, lets add an additional column to matrix $A$

```{r}
P = matrix(c(30,60,90),3,1)
print(P)
```
```{r}
Q = cbind(A,P)
print(Q)
```

### Matrix Algebra

In Matrix algebra, we have

- Matrix Addition
- Matrix Subtraction
- Matrix Multiplication (Regular)
- Matrix Multiplication (Elementwise)
- Matrix Division (Elementwise)


Let's create 2 sample matrix of $(3\times 3)$ order to perform such operations :

```{r}

mat1 = matrix(c(1,2,3,4,5,6,8,9,1),3,3, byrow = T)
mat2 = matrix(c(3,1,3,4,2,1,5,1,2),3,3, byrow = T)

print(mat1)
```
```{r}
print(mat2)
```

1. Matrix Addition :

```{r}
mat1 + mat2
```

2. Matrix Subtraction :

```{r}
mat1 - mat2
```

3. Matrix Multiplication (Regular) :

```{r}
mat1 %*% mat2
```

4. Matrix Multiplication (Elementwise) :

```{r}
mat1 * mat2
```

5. Matrix Division (Elementwise) :

```{r}
mat1 / mat2
```












