---
title: "Recasting With `reshape2` Library"
output: html_document
---
Using the `reshape2` package, we can perform the recasting operation on the dataframe.

**Recasting** is a process of manipulating a dataframe in terms of its variable.

Recasting has two steps :

  - Melting
  - Casting

The types of variable used for recasting are :

  - Identifier Variable
  - Measurement Variable
  
### Identifier Variable

Discrete variables like "Name", "Month" etc.

### Measurement Variables

Numeric variables like "Weight", "Age" etc.

***Note :*** *Categorical and date variable are not measurement variables.*

Let's create a dataframe for our illustration :

```{r}
Name = c("Senthil", "Senthil", "Sam", "Sam")
Month = c("Jan", "Feb", "Jan", "Feb")
BS = c(141.2, 139.3, 135.2,160.1)
BP = c(90, 78, 80, 81)

ed = data.frame(Name, Month, BS, BP)
print(ed)
```

### `reshape2` Installation For Recasting Operation

For recasting operation, we need a R-library named `reshape2`.

We have to write the following code for installing and calling the library into our console :

```{r, eval = FALSE}
install.packages("reshape2") #For installing the package
library("reshape2") #For calling the package
````


### The Melt Operation :

The following code perfroms the "melt" operation :

```{r}
library("reshape2")

DF = melt( ed, 
           id.vars = c("Name","Month"), 
           measure.vars = c("BS","BP") )
print(DF)
```

- By `id.vars` we have loaded our identifier variables.
- By `Measure.vars` We have loaded our measurement variables.
- The melt operation basically unpivots the measurement variables.

### The Cast Operation :

If we have two columns with categorical variables named as `col1` And `col2` then,

`col1 ~ col2` Means we want to show `col1` by `col2`.So, `col1` will remain as it is but, each element of `col2` has to become variables/columns to show the respective values of `col1`.
  
Let's see an example of casting operation as follows :

```{r}
library("reshape2")
DF2 = dcast(DF, 
            variable+Month ~ Name, 
            value.var = "value")
print(DF2)
```

- `variable+Month ~ Name` means we want to see `Variable` and `Month` by `Names`. So, the attributes of the `Names` column will become separate variables.
- `Value.var` arguments takes the entity(value) by which we want to show the attributes of `Names` column against the `Variable` and `Month`.
- The cast operation pivots an identifier variable.

### Recasting Operation In Single Step

We can perform both melting and casting operations on the dataframe `ed` in a single line as follows :

```{r}
library("reshape2")
DF3 = recast(ed,
             variable+Month ~ Name, 
             id.var = c("Name","Month"))

print(DF3)
```

- The first argument takes the dataframe `ed` to work upon.
- The second argument performs the casting operation.
- The third argument performs the melting operation by indicating the identity variables to the machine with `id.var` command.
- The machine considers the rest variables as measurement variables once we define our identity variables.
