---
title: "Basic Distributions"
---
Some of the basic statistical distributions are :

1. Normal Distribution
2. Binomial Distribution
3. Poisson Distribution
4. t Distribution
5. F Distribution
6. Exponential Distribution
7. Chi Squared Distribution

To get help on various distributions available the `stats` package of R, pass the following command :

```{r disthelp, eval=FALSE}
help("distributions") #or,
?distributions
```

### Normal Distribution

Let's say,a variable ***X*** is normally distributed with a known ***mean = 75*** and ***Standard Deviation (SD) = 5***.

So,

`X ~ NORM(mu = 75, var = 25)

To get some help on normal distribution in R :

```{r nomhelp, eval=FALSE}
help("pnorm")
#or,
?pnorm
```

To find the probability of ***X <= 70***, i.e., `P(X <= 70)` is :

```{r norm}
pnorm(q=70, mean = 75, sd = 5, lower.tail = TRUE )
```
- In R, `lower.tail=TRUE` is default argument.

To find the probability of ***X >= 85***, i.e., `P(X >= 85)` is :

```{r norm2}
pnorm(q=85, mean = 75, sd = 5, lower.tail = FALSE )
```
The `pnorm()` command can also be used to calculate the probability when ***z-Value*** is given.

To find the probability of ***z >= 1***, i.e., `P(z >= 1)` is :

```{r norm3}
pnorm(q=1, mean = 0, sd = 1, lower.tail = FALSE )
```
To find the 25th percentile (1st quartile), we can use the `qnorm()` command :

```{r qnorm}
qnorm(p = 0.25, mean = 75, sd = 5, lower.tail = TRUE)

```

Let's plot a probability density function :

```{r sequence}

x = seq(from = 55, to = 95, by = 0.25)
x

```

Now, let's calculate the propability density for each element of list `x`.

```{r dnorm}
dens = dnorm(x, mean = 75, sd = 5)
plot(x, dens, 
     type = "h",
     xlab = "x",
     ylab = "Probability Density",
     las = 1,
     main = "Normal Distribution",
     col = "dark red"
     )

abline(v = 75, col = "red", lwd = 2)
```

To get a random sample from a normally distributed population, we can use, the `rnorm()` command.

```{r rnorm}
randsample = rnorm(n=40, mean = 75, sd = 5)
randsample
```
Let's quickly make a histogram to see the distribution of the sample :

````{r rnormdist}

hist(randsample,
     xlab = "Sample",
     ylab = "Frequency",
     main = "Sample Distribution",
     col = "seagreen",
     border = "white")

```

From the above histogram, we can see that even the distribution of random sample is not normally distributed even though it has been taken from a normally distributed population.


### Binomial Distribution

Let's say,a variable ***X*** is binomially distributed with ***n=20*** trials and ***p=1/6*** probability of success.

So,

`X ~ BIN(n = 20, p = 1/6)

To get some help on binomial distribution in R :

```{r binomhelp, eval=FALSE}
help("dbinom")
#or,
?dbinom
```

- The `dbinom` command is used to find the values for the probability density function of ***X, f(X)***


So, to find the probability of ***X=3***, i.e., `P(X=3)` :

```{r dbinom}
dbinom(x=3, size = 20, prob = 1/6)
```

So , the probability of exactly 3 successs on 20 trials is 23.7% (approx.)

To find the probability of ***X=0 & X=1 & X=2 & X=3***, i.e., `P(X=0) & P(X=1) & P(X=2) & P(X=3) ` is :
 
```{r dbinom2}
dbinom(x=0:3, size = 20, prob = 1/6)
```
So, we can see :

- Approx. 2.6% of getting 0 success in 20 trails
- Approx. 10.4% of getting 1 success in 20 trails
- Approx. 19.8% of getting 2 success in 20 trails
- Approx. 23.7% of getting 3 success in 20 trails

Let's plot a probability density function :

```{r binomsequence}

binomlist = seq(from = 0, to = 150, by = 1)
binomlist

```

Now, let's calculate the propability density for each element of list `binomlist`.

```{r dbidist}
binomdens = dbinom(binomlist, size = 150, prob = 1/2)
plot(binomlist, binomdens, 
     type = "h",
     xlab = "x",
     ylab = "Probability Density",
     las = 1,
     main = "Binomial Distribution",
     col = "dark red"
     )
abline(v = 75, col = "red", lwd = 2)

```

To find the probability of ***X <= 6***, i.e., `P(X <= 6)` is :

```{r dbinom3a}
sum(dbinom(x=0:3, size = 20, prob = 1/6))
```
or,
we can use the `pbinom()` command to get the probability distribution function of ***X, f(X)***
```{r pbinom}
pbinom(q=3, size = 20, prob = 1/6, lower.tail = T)
```

So, the probablity of getting 3 or, fewer success in 20 trials is approx. 56.6%.

- The `rbinom()` command is used to take a random sample from a binomial distribution.
- The `qbinom()` command is used to find quantiles for a binomial distribution.

### Poisson Distribution

Let's say,a variable ***X*** follows a poisson distribution with a known rate of ***lamda=20***

So,

`X ~ POISSON(lambda = 7)

To get some help on Poisson distribution in R :

```{r poissonhelp, eval=FALSE}
help("dpois") #or,
?dpois
```

- We can calculate the probabilities for Poisson distribution using the `ppois()` and `dpois()` commands

To find the probability of ***X=4***, i.e., `P(X=3)` :

```{r dpois}
dpois(x = 4, lambda = 7)
```

So, there is approximately 9% chance of exactly 4 occurances.

To find the probability of exactly ***X=0 & X=1 & X=2 &...& X=4***, i.e., `P(X=0) & P(X=1) & P(X=2) &...& P(X=4) ` is :

```{r dpois2}
dpois(x = 0:4, lambda = 7)
```
So, we can see :

- Approx. 0.09% chance of exactly 0 occurances.
- Approx. 0.6% chance of exactly 1 occurances.
- Approx. 2.2% chance of exactly 2 occurances.
- Approx. 5.2% chance of exactly 2 occurances.
- Approx. 9.1% chance of exactly 4 occurances.

Let's plot a probability density function :

```{r poissonsequence}

poislist = seq(from = 0, to = 200, by = 0.5)
poislist

```

Now, let's calculate the propability density for each element of list `binomlist`.

```{r dbidist}
poisdens = dpois(poislist, lambda = 50)
plot(poislist, poisdens, 
     type = "h",
     xlab = "x",
     ylab = "Probability Density",
     las = 1,
     main = "Poisson Distribution",
     col = "dark red"
     )
abline(v = 50, col = "red", lwd = 2)

```

To find the probability of ***X <= 4***, i.e., `P(X <= 4)` is :

```{r dpois3}
sum(dpois(x = 0:4, lambda = 7))
```
or,
we can use the `ppois()` command to get the probability distribution function of ***X, f(X)***
```{r ppois1}
ppois(q=4, lambda = 7, lower.tail = T)
```

So, there is approximately 17.2% chance of exactly 4 or, fewer occurances.

Similarly, to find the probability of ***X >= 12***, i.e., `P(X >= 12)` is :

```{r ppois2}
ppois(q=12, lambda = 7, lower.tail = F)
```

So, there is approximately 2.6% chance of exactly 12 or, more occurances.

- The `rpois()` command is used to take a random sample from a Poisson distribution.
- The `qpois()` command is used to find quantiles for a Poisson distribution.

### t Distribution

## Getting Brief About **t-Distribution**

To get help of *t-distribution*, we can simple write :

```{r help option-1, eval=FALSE}
help(pt)
```
or,

```{r help option-2, eval=FALSE}
?pt
```

Let's say we are given :

*t-statistics* = 2.3
*Sample Size (n)* = 26

So, 

*Degrees of freedom (df)* = n-1 = 25

To get *p(t > 2.3)* :

```{r p-value from t-stat}
pt(q = 2.3, df = 25, lower.tail = FALSE)
```
So, by the argument `lower.tail = FALSE`, we mean we want to find the the area after `t > 2.3`.

#### Getting **Two sided p-value** :

This is used to get the area above `t = 2.3` and below `t = -2.3`

To get this :

```{r two sided p-value from t-stat}
pt(q=2.3, df=25, lower.tail = F) + pt(q=-2.3, df=25,lower.tail=T)
```

Or,

```{r two sided p-value from t-stat opt-2}
pt(q=2.3, df=25, lower.tail = F) * 2
```

### Calculating **t-value** for 95% Confidence :

95% confidence implies the *p-value* in each tail as 2.5%

So, to find *t-statistics* when *p-value of each tail is 2.5%* :

```{r t-value for 95% conf.}
qt(p = 0.025, df = 25, lower.tail = T)
```


