---
title: "Statistics Fundamentals"
---

It would not be wrong if we say that statistics starts with an interest to quantify the centre and spread of the distribution of a variable.

So, let's load the "Lung Capacity Dataset" to produce the numerical summeries of categorical and numerical variables.

```{r dataload}
LungCapData = read.table(file = "Dataset/LungCapData.txt", header = TRUE, sep = "\t")
head(LungCapData)
attach(LungCapData)
```
The above table gives us a brief look over the data andd class of each variable.

Now, let's see various summarizing options for the categorical and numerical variables


### Summarizing Categorical Variables


Categorical variables can be summarized using a frequency/proportion table and to produce a frequency table, we use either use the `table()` function or, the `summary()` function :

```{r CategoricalSummary}
table(Smoke)
```
Or,

```{r CategoricalSummary2}
summary(Smoke)
```

So, we can see, we have `648` observations for `no` and `77` observations for `yes` within the categorical variable `Smoke`.

To get the total number of observations in a variable, we use the `length()` command :

```{r LengthOfVariable}
length(Smoke)
```

Now, we can produce a proportion table for `Smoke` variable, as follows :

````{r ProportionTable}
table(Smoke)/length(Smoke)
```

To produce a two-way table or, contengency table between `Gender` and `Smoke`, we can utilise the `table()` as follows :

```{r ContengencyTable}
table(Smoke, Gender)
```

***Note :*** The first variable passed in the `table()` function is taken into the rows.

### Summarizing Numeric Variables

We can get the mean of our numeric variable `LungCap` with `mean()` function :

```{r mean}
mean(LungCap)
```
To get a trimmed mean :

```{r meantrim}
mean(LungCap, trim = 0.10)
```
***Note :*** By the argument `trim = 0.10`, we have removed the top and bottom 10% observations from the mean calculation.

To calculate median, we can use the `median()` command :

```{r median}
median(LungCap)
```
To calculate the variance, we use the `var()` command :

```{r variance}
var(LungCap)
```

To calculate the standard deviation, we can the `sd()` command or, can simply pass the `var()` argument within th `sqrt()` function :

```{r stdev}
sd(LungCap)
```
Or, 

```{r stdev2}
sqrt(var(LungCap))
```

similarly, the alternate way to calculate the variance is to sqaure the standard deviation :

```{r Variance2}
Variance = sd(LungCap) ^ 2
Variance
```

To get the minimum and maximum values present in the `LungCap` variable :

```{r minimum}
#For Minimum
min(LungCap)
```
and 
```{r maximum}
#For Maximum
max(LungCap)
```

We can find the range(minimum & maximum) of the variable, by using the `range()` command :

```{r range}
range(LungCap)
```

Specific quantiles/percentiles can be calculated for the variable, using the `quantile()` function :

To get the 90th percentile :

```{r quantile}
quantile(LungCap, probs = 0.90)
```
To obtain multiple percentiles, we can pass a list in the `probs` argument :

```{r quantiles2}
quantile(LungCap, probs = c(0.2, 0.5, 0.7, 1))
```

To calculate the sum of all the entries of a variable :

```{r sum}
sum(LungCap)
```

To calculate the Pearson's correlation between `Age` and `LungCap`, we can use the `cor()` function:

```{r corr}
cor(LungCap, Age)
```

***Note :*** The default correlation in `cor()` function is Pearson's correlation but, if we need any other type of correlation, then, we have to pass an additional argument `method`.

To get Spearman's correlation between `Age` and `LungCap`

```{r CorrSpearman}
cor(LungCap, Age, method = "spearman")
```

The covariance between `Age` and `LungCap` can be calculated using `cov()` command or, by `var()` command as well.

```{r cov1}
cov(LungCap, Age)
```
Or,

```{r cov2}
var(LungCap, Age)
```

We can use the `summary` command, to get most of the summaries for a variable :


```{r summary}
summary(LungCap)
```

As we can see, the output of `summary()` function gives us the following summary level information of the variable :

1. Minimum
2. 1st quartile (25th percentile)
3. Median (50th Percentile/2nd quartile)
4. Mean
5. 3rd quartile (75th percentile)
6. Maximum

To get help in any of the summary attributes, just pass its name inside the `help()` command or, after `?`, as follows :

```{r helpsummary, eval=FALSE}
help("mean") #or,
?mean
```

We can also pass the entire dataset inside the `summary()` command to get a cursory look over the summary of each variable :

```{r tablesummary}
summary(LungCapData)
```

















