---
title: "t-Test"
---


### One Sample t-Test

To perfrom a one sample t-Test, let's first get the data :

```{r Getting the data}
LungCapData = read.table(file = "Dataset/LungCapData.txt", header = T, sep = "\t")
```

Let's see the variables present in our dataset :

```{r view the headers}
names(LungCapData)
```

```{r checking Class}
class(LungCapData$LungCap)
```

To get any additional help on *t-test* :

```{r help, eval=FALSE}
help("t.test")
#or,
?t.test
```

Let's visualize the data first :

```{r boxplot/}
boxplot(LungCapData$LungCap)
```

Let's say,

Ho:  mean = 8
Ha: mean < 8

Let's perform the hypothesis testing

```{r hypothesis test}
t.test(LungCapData$LungCap, mu=8, alternative = "less", conf.level = 0.95)
```
As the p-value is more than 5% so, we failed to reject the Ho and hence, the sample mean is less than 8.

The default t-test in R is two sided, so, if we not include the argument `alternative="less"` then, R will perform a two sided hypothesis test.

To see the attributes of the t-Test, we can store the test results in a variable and then pass the `attributes` function :

```{r test attributes}
Test = t.test(LungCapData$LungCap, mu=8, alternative = "less", conf.level = 0.95)
attributes(Test)
```

To only see a specific attribute of the test :

```{r t-test conf. int}
Test$conf.int
```

```{r t-test p-value}
Test$p.value
```

### Paired T-Test

Let's first get the data :

```{r BP data}

BPData = read.table(file = "Dataset/BloodPressure.txt",header = T, sep = "\t")

```

Let's see the data:

```{r eda}
names(BPData)
```
dimesion of the dataset is :

```{r size}
dim(BPData)
```

Let's visualize the data :

```{r boxplot}
attach(BPData)
boxplot(Before,After)
```

Let's plot a scatter plot between the "Before" and "After" result to see a correlation between them :

```{r correlation}
plot(Before, After)
abline(a=0, b=1) #slope = 0 and intercept = 1
```

Let's perform a two sided hypothesis test on our data :

Ho: Mean difference is zero, i.e., no chnage is SBP before and after
Ha: Mean difference is not zero, i.e., there is a change is SBP

```{r paired t-test}
t.test(Before, After, mu=0, alternative = "two.sided", paired = TRUE, conf.level = 0.99)
```
As the *p-Value* is very small, hence we reject the null hypothesis.

If we change the order of "Before" and "After" then, it will only change the sign of `mean if the differences`, `Confidence Interval` and `t-Statistics`.

### Independent Two Sample t-Test

- A parametric method appropriate for examining the difference in means for 2 populations.

- A way of examining the relationship between a numeric outcome variable (Y) and a categorical explanatory variable (X, with 2 levels)

We will be examining the relationship between the **Smoking Habits** and its effect on **Lung Capacity**.

Let's see the class of these two variables first.

```{r checking the class}
class(LungCapData$LungCap)
class(LungCapData$Smoke)
```
Let's visualize the data :

```{r boxplot Lungcap~Smoke}
attach(LungCapData)
boxplot(LungCap ~ Smoke)
```

Let's perform a hypothesis test :

Ho: Mean Lung capcity of smokers and non-smokers are same
Ha: Mean Lung capacity of smokers and non-smokers are different.

Let's perform a two sided t-test -:

```{r two sided t-test}
t.test(LungCap ~ Smoke, mu=0, alt="two.sided", conf=0.95, var.eq=F, Paired=F )
```

As the **p-Value** is small, so we reject the null hypothesis and thus, the mean lung capacity of smokers and non-smokers are different.

By the argument `var.eq=F`, we considered non-equal variance for the groups.

```{r variance check}

var(LungCap[Smoke=="yes"])
var(LungCap[Smoke=="no"])
```
We can clearly see that the variance in lung capacity for people who doesn't smoke is almost double of those who smoke.

### Levene's Test

This test is used to check whether the population variances are equal or, not.

So, let's perform an hypothesis test on this.

Ho: Population variances are equal
Ha: population variances are not equal

```{r loading package to perform test}
library(car)
```

Let's now perform the **Levene's Test**

```{r Levene test}
leveneTest(LungCap~Smoke)
```

From the above test, we can see :

***p-Value :*** 0.00034 = 0.034%

As the *p-Value* is very small, so, we reject the null hypothesis and have evidence to believe that the population have unequal variances.














